---
title: "Data Wrangling, EDA, and Baseline Model"
authors: "Dasha Metropolitansky, Ilkin Bayramli, Chang Yu, Maarten De Vries"
date: "21/9/2020"
output: pdf_document
---

# Variable Selection & Data Wrangling
```{r include=FALSE, message=F, warning=F}
# Loading packages
library(tidyverse)
library(janitor)
library(countrycode)
```

```{r include=FALSE, message=F, warning=F, cache=T}
knitr::opts_chunk$set(echo = TRUE)

# Reading in the data
load('WVS_Cross-National_Wave_7_R_v1_4.rdata') 
df_full = `WVS_Cross-National_Wave_7_R_v1_4` %>% 
  clean_names()

rm(`WVS_Cross-National_Wave_7_R_v1_4`)
```

## Filtering for observations from the 7th wave of the World Values Survey
```{r message=F, warning=F, cache=T}
df_full = df_full %>% 
  filter(a_study == 2 & a_wave == 7)
```

## Selecting variables
```{r message=F, warning=F, cache=T}
# Selecting relevant variables
df_filtered = df_full %>% 
  select(a_year, b_country_alpha, g_townsize2, h_settlement, h_urbrural,
         q46, q47,
         q94, q95, q96, q97, q98, q99, q100, q101, q102, q103, q104,
         q173,
         q209, q210, q211, q212, q221, q222,
         q260, q262, q263, q264, q265, q269, q270, q271, q273, q274, 
         q275r, q277r, q278r, q279, q281, q284, q285, q288r, q289)

# Changing variable names
colnames(df_filtered) = c('year', 'country', 'town_size', 'settlement_type', 'urban',
                 'happy', 'healthy',
                 'church_member', 'sport_member', 'arts_member', 'union_member', 'political_party_member',
                 'environment_member', 'prof_member', 'charity_member', 'consumer_member', 'self_help_member',
                 'women_member', 
                 'religious', 
                 'political_1', 'political_2', 'political_3', 'political_4', 'political_5', 'political_6',
                 'gender', 'age', 'immigrant', 'immigrant_mother', 'immigrant_father', 'citizen', 'household_size',
                 'live_with_parents', 'relationship', 'num_kids', 'education', 'education_mother',
                 'education_father','employment_status', 'occupation', 'sector', 'breadwinner', 'income', 'religion')

happy_full = df_filtered$happy
```

## Modifying variables
```{r message=F, warning=F, cache=T}
df = df_filtered %>%
  # Converting urban into an indicator variable
  mutate(urban = case_when(urban == 1 ~ 1,
                           urban == 2 ~ 0,
                           TRUE ~ NA_real_),
  # Converting happy into an indicator variable
         happy = case_when(happy == 1 | happy == 2 ~ 1,
                           happy == 3 | happy == 4 ~ 0,
                           TRUE ~ NA_real_),
  # Converting health into an indicator variable
         healthy = case_when(healthy %in% c(1,2,3) ~ 1,
                             healthy %in% c(4,5) ~ 0,
                             TRUE ~ NA_real_)) %>% 
  # Converting the membership variables into indicator variables
  mutate_at(
    vars(one_of('church_member', 'sport_member', 'arts_member', 'union_member', 'political_party_member',
                 'environment_member', 'prof_member', 'charity_member', 'consumer_member', 'self_help_member',
                 'women_member')),
    list(~case_when(
      . == 2 ~ 1,
      . == 1 | . == 0 ~ 0,
      TRUE ~ NA_real_))) %>%
  # Converting religious into an indicator variable
  mutate(religious = case_when(religious == 1 ~ 1,
                               religious == 2 | religious == 3 ~ 0,
                               TRUE ~ NA_real_)) %>%
  # Assigning responses to the political variables a score 
  mutate_at(
    vars(one_of('political_1', 'political_2', 'political_3', 'political_4', 'political_5', 'political_6')),
    list(~case_when(
      . == 1 ~ 2,
      . == 2 ~ 1,
      . == 3 ~ 0,
      TRUE ~ NA_real_))) %>%
  # Creating one political variable which is the sum of political variables' scores 
  # mutate(political = rowSums(select(., political_1:political_4))) %>%
  # select(-c(political_1:political_6)) %>%
  # Converting gender into an indicator variables
  mutate(male = case_when(gender == 1 ~ 1,
                          gender == 2 ~ 0,
                          TRUE ~ NA_real_)) %>% 
  select(-c(gender)) %>%
  # Converting immigration variables into indicator variables
  mutate_at(
    vars(one_of('immigrant', 'immigrant_mother', 'immigrant_father')),
    list(~case_when(
      . == 2 ~ 1,
      . == 1 | . == 0 ~ 0,
      TRUE ~ NA_real_))) %>% 
  # Converting citizen into an indicator variable
  mutate(citizen = case_when(citizen == 1 ~ 1,
                             citizen == 2 | citizen == 3 ~ 0,
                             TRUE ~ NA_real_),
  # Converting live_with_parents into an indicator variable
         live_with_parents = case_when(live_with_parents == 1 ~ 0,
                             live_with_parents %in% c(2,3,4) ~ 1,
                             TRUE ~ NA_real_),
  # Converting relationship into an indicator variable
         relationship = as.factor(relationship),
  # Converting breadwinner into an indicator variable
         breadwinner = case_when(breadwinner == 1 ~ 1,
                                 breadwinner == 2 ~ 0)) %>%
  # Replacing all entries which are negative numbers with NA
  mutate_each(funs(replace(., .<0, NA))) %>% 
  # Converting relevant variables to factors
  mutate_at(
    vars(one_of('country', 'town_size', 'settlement_type', 'urban', 'happy', 'healthy',
                'church_member', 'sport_member', 'arts_member', 'union_member', 'political_party_member',
                 'environment_member', 'prof_member', 'charity_member', 'consumer_member', 'self_help_member',
                 'women_member', 'religious', 'male', 'immigrant', 'immigrant_mother', 'immigrant_father',
                'citizen', 'live_with_parents', 'relationship', 'education', 'education_mother', 'education_father',
                'employment_status', 'occupation', 'sector', 'breadwinner', 'income', 'religion', 'political_1',
                'political_2', 'political_3', 'political_4', 'political_5', 'political_6')),
    list(~as.factor(.)))

# Removing unnecessary data frames
rm(df_full, df_filtered)

# Call garbage collector
gc()
```

# EDA

This dataset tries to cover political, social, economic, spiritual, geographic aspects of the survey participants which can be related to their happiness levels. 
The dataset also includes geographic variables that can affect the happiness of the survey participants. `country` of the participant is one such variable which includes the following countries:

```{r message=F, warning=F}
unique(df$country)
```

Variables such as `urban, settlement_type, town_size` further refine the geographic details pertaining to the residence of the surveyees. The dataset also has some variables which deal with the legal status of the participant in the country of the survey as well as their / their parents' immigration status. 

Economic conditions of the participants are covered by variables such as `income` bracket (low/med/high), `household_size`, `employment_status`, `breadwinner`. The variable `occupation` describes 12 categories of jobs the participant works in. The political views of the survey takers are most prominently accounted for by the categorical variables `political_party_member` and cumulative numeric predictors `ideology`, `political` which add up different correlated binary variables to represent political activeness and right-wing ideology of a given participant respectively. These additional variables describe the membership of a participant in various economic/political/social organizations:

```{r message=F, warning=F}
colnames(df)[grepl("member", colnames(df))]
```

The variable `male` describes the gender of the study participant, and about `r sum(df$male[!is.na(df$male)] == 1) / length(df$male[!is.na(df$male)])` portion of surveyees are male.

Although the majority of variables are categorical, there are some numerical variables such as `num_kids, age, household_size` and some cumulative scores we designed such as `political, ideology`.`Age` is another numeric variable worth looking which follows this distribution:

```{r warning=FALSE,message=FALSE, fig.width=15, fig.height=7,fig.align='center'}
ggplot(data = df, aes(x = age)) + 
  geom_histogram(color="black", fill = "firebrick4")
```

As seen the study covers participants from almost all age groups with ages 15 - 40 being the most prominently covered. 

The only temporal variable we have is `year` which ranges from `r min(df$year)` to `r max(df$year)`. Here is the mean happiness and number of survey participants by year who did not have `NA` value for their happiness score:

```{r warning=FALSE,message=FALSE, fig.width=15, fig.height=7,fig.align='center'}
df %>%
  select(year, happy) %>% 
  filter(!is.na(happy)) %>% 
  mutate(happy = ifelse(happy == 1, 1, 0)) %>%
  group_by(year) %>%
  summarize(mean_happy = mean(happy), n = n())
```

Finally and most importantly, here is the distribution of the response variable `happy`:

```{r warning=FALSE,message=FALSE, fig.width=15, fig.height=7,fig.align='center'}
ggplot(data = tibble(happiness_level = happy_full), aes(x = happiness_level)) + 
  geom_histogram(color="black", fill = "orange")
```

Where the happiness levels translate into:
```
4 = Very happy
3 = Rather happy
2 = Not very happy
1 = Not at all happy
```
Since the different between `3, 4` and `1, 2` are rather vague (and per recommendation of a TF), we decided to treat this variable as binary for now with `3, 4` being `1` and `1, 2` being a `0`.

# Baseline Model

We wanted to start by fitting a full linear model for the baseline. One issue is that our predictor variable is categorical which linear regression is poorly fit for (and not covered in this class so far). We will try to address this issue in future models.

```{r warning=FALSE,message=FALSE, fig.width=15, fig.height=7,fig.align='center'}
df$happy_cont = ifelse(df$happy == 1, 1, 0)
lm1 = lm(happy_cont ~ ., data = df)
summary(lm1)
```

Out of this model, 4 very significant predictors of happiness arise: `country`, `citizen`, `happy`, and `town_size`. The significance of these variables is quite intuitive, however, some more significant variables are likely to be uncovered once we fix assumptions of violations (our response is binary!). 
